---
title: "Wrapper Functions Basic Writeup"
bibliography: references.bib
format:
  pdf:
    pdf-engine: pdflatex
editor_options:
  chunk_output_type: console
header-includes:
  - \usepackage{fancyvrb}
  - \DefineVerbatimEnvironment{fadecode}{Verbatim}{fontsize=\small, frame=single, commandchars=\\\[\]}
---

# Abstract

Wrapper functions are a common class of functions that convert many lines of code into a function
with a few inputs for key modifications.
Wrappers can greatly simplify using complex code, making it far more accessible to less experienced
users.
However, many possible modifications to the original code may become inaccessible, making edits not
foreseen by the wrapper creator difficult to impossible for users.
There are a few ways for wrapper creators to make future edits to their function easier.
These start with the function being self-contained
(\textcolor{magenta}{Do I put a definition for self-contained here? Or is it sufficient it is defined
when brought up in the body of the paper?}), but also include a mechanism for users to view and edit
code, and potentially even save their modified versions of the wrapper function for future use. 

# What are Wrapper functions

## Basic description

Wrapper functions are a broad category of functions that typically aim to take a more complex task
and create a simple way of recreating it, often by limiting the number of required inputs
(\textcolor{magenta}{This is pretty broad, should I focus on a more narrow category of wrapper functions.
I kind of bounce back and forth between ones that generate figures and ones that generate models}).
To accomplish this, a more complex block of code is wrapped up in a function and many inputs have
values selected by the creator.
A few key inputs are left outward-facing to be adjusted by the user with the wrapper function.
However, many inputs may be permanently unadjustable by the user.
Common uses for wrapper functions include complex graphics where there may be hundreds of lines of
code, but only a couple are typically of interest to the user.
Another common use is in modeling functions where a single function call may test a great number of
models, assumptions about the data, clean the data, and return what it considers to be the
best-fitting model with very little user input.
(\textcolor{magenta}{Are there other major uses I'm blanking on})

## Popular examples

A popular class of wrapper functions helps users create complex ggplots involving multiple layers
with as little as one line of code [@wickham2016ggplot2].
A major example of this style is `autoplot()`, which uses the class of the object given to it to
choose an appropriate ggplot display.
This allows users to quickly plot objects like time series, model objects, or dimension reduction
results without needing to know all of the individual ggplot layers being created.
The **survminer** package provides a more domain-specific example with `ggsurvplot()`, a wrapper
that builds survival curves from fitted survival objects and can add common survival-analysis
features such as confidence intervals, censoring marks, p-values, and risk tables
[@kassambara2026survminer].
The **ggpubr** package takes a broader plotting-helper approach, providing functions such as
`ggboxplot()` and `ggscatter()` that wrap common ggplot patterns behind simple arguments for labels,
grouping, colors, and statistical annotations [@kassambara2026ggpubr].
There are also packages created as companions for textbooks that often have wrapper functions to
help create graphics like those seen in the book.
The textbook **Time Series Analysis and Its Applications** [@shumway2025time] has its own package
**astsa** with several wrappers for creating similar graphics, and for quickly determining models
[@stoffer2025astsa].
Even packages not focused on producing graphics may have one or two wrappers for creating specific
visuals related to the goal of their package.

Another important class of wrapper functions is involved in modeling.
The **fable** package in time series has modeling functions that can take in as little as just a
dataset and from there the function will make numerous decisions for the user to return what it
determines to be the best model for their data [@oharawild2024fable].
(\textcolor{magenta}{Do I provide more modeling examples, do i provide more details.})

## Advantages

The clear advantage to creating a wrapper function is enabling users to create complex figures or
well-fitted models with ease.
For a user looking to complete a single application with as little friction as possible this is ideal.
With a wrapper function the creator can enable beginner users to recreate their more advanced figures
or models.
This may be important if the user has colleagues that need to replicate work in a similar style,
since it reduces the expertise needed by the user, and increases consistency.
If none of the inputs have been fixed and made inaccessible to the user, then wrapper functions offer
primarily upside.
In some cases what would have been hundreds of lines of code becomes one.

Outside of mere ease of use, wrapper functions can allow a user to lean on the experience of the
creator of the function.
If the creator is an expert in the area, their preselected values for inputs, or methods for
automatically determining them, may be more reasonable than anything an inexperienced user could
guess.
The user is less likely to violate unknown assumptions if the function checks those itself.
Core code may have even been converted into C or other faster running languages, making the function
faster than anything the user could easily write themselves.

## Issues

Issues can begin to crop up if an input that was fixed by the creator is something you need to
adjust.
It could be as simple as switching the line colors of a plot to be more colorblind friendly.
Depending on how the wrapper function was written, making minor adjustments can range from a slight
inconvenience to a total nightmare as you try to navigate the depths of their package architecture
searching for where the color of those lines was actually encoded.
The simplest wrapper function is just a block of code that has been placed in a function.
We call these styles of wrapper functions "self-contained" since the function itself contains
everything it needs to function.
Self-contained wrappers are very easy to adjust since you can take the block of code and create a new
wrapper with the code slightly adjusted, or by just adding a new input to allow yourself to adjust
what was previously fixed.
Difficulties arise when the function is not self-contained.
The wrapper may call many other functions, each of which could conceivably contain the code you need
to adjust, requiring you to search through all of them.
Outside of just calling other functions, the function could use S3, S4, or R6, giving it custom
methods that could further adjust how the output looks.
If the function calls compiled C code that contains what you want to adjust, there may be nothing you
can do outside of writing it all yourself or finding a new package.

Beyond difficulty with modifying the function, wrapper functions can be very difficult to bug fix.
The resulting error when something goes wrong can be very unclear, particularly if the error is in
another function called by the wrapper function.
The user often struggles to determine if the issue is with their inputs, or if they are using the
wrapper function incorrectly.
The error may be fatally incomprehensible if the error is with a compiled C function.

Finally, if a user would like to use the output of the wrapper as part of some greater project, they
may be unsure what they have actually done.
(\textcolor{magenta}{There's plenty to be said about the user not really understanding their code
and what they have really done, but is that something worth going down. My proposals in this paper
don't really do anything to fix that issue.})

# Ways to combat Basic Issues

Package creators typically create wrapper functions to make their code as usable for users as
possible.
So, what steps exist to allow wrapper functions to be as simple and functional, but open to
adjustment as possible?
One unfortunate drawback to all the following methods is that they work better the more
self-contained the function is.

## Original method

One adjustment to allow users to make simple adjustments to a function could be to have an argument
that by default runs the wrapper as usual, but could be switched to print out the code to be run in a
new document instead of running the function.
Adjustments to the code could be made live before running it.
This would be helpful for very minor changes like switching the color of a line in a plot, or
changing a couple fixed numeric values.
While this method is easy to implement and nice for single adjustments, it is inconvenient if a user
would like to consistently make the same change when using the function a number of times.

## Binding method

While the method outlined in section a is nice for minor one-off changes, it would be rather
inconvenient if an adjustment was desired to be used regularly.
This is especially true if you want to use the wrapper function many times, all with the same
adjustment.
To build off the previous method, you could instead bind the adjusted code to a new object in the
working environment.
The wrapper function could be adjusted to take in modified code and run it instead of its original.
The user could then easily make their adjustment one time and continue to use their modified version
of the original wrapper function.
The wrapper function could even allow the name of the binding to be passed into the function itself
and then assign it to the global environment after adjustments have been made.
This creates a more semi-permanent solution, but the edited code living in the global environment
requires the user to re-create it, or save it as an RDS, between sessions.
It also requires the creator to add several arguments to every wrapper function they would like to be
editable by the user.

## Template method

The previous methods have both involved editing wrapper functions to have their typical functions as
well as arguments that give them editing functions.
For a single wrapper this can be useful, but for a package containing multiple it may be cleaner to
have a single editing function that works for all wrapper functions.
In this method the code bodies of wrapper functions are stored in the package, and a dedicated
templating function calls them into a new document.
An argument for the template's name is given, and when edits to the function are complete the edited
code is saved as an RDS, as opposed to just being in the global environment.
These templates allow for long-term use of modified code for the user, and even for templates to be
shared between users if several would like to make the same adjustment.
The only adjustment that needs to be made to the wrapper functions themselves in this method is
allowing for a template to be provided and running the template code instead.

# How to Implement

To give a running example of how this might work, we've created a wrapper function called raincloud
that takes in a tibble with two variables, value and class.
The default wrapper produces a violin plot with jittered observations, a narrow boxplot, and a median
point. 

```{r, echo=FALSE}
pkgload::load_all(".", quiet = TRUE)
```


```{r}
#| label: raincloud-data
#| message: false
#| warning: false
set.seed(17)
a <- tibble::tibble(
  value = c(rnorm(2e2, 1), 4 * rbeta(2e2, 1, 4)),
  class = c(rep("a", 2e2), rep("b", 2e2))
)
```

```{r}
#| label: raincloud-default
#| fig-cap: "Default raincloud wrapper output."
raincloud_original(x = a)
```

The following methods also utilize copies of each function being stored in the extdata folder of the package so that they can easily be read for editing. 

## Original method

For the original method, the package creator first adds an argument such as `format` to the wrapper. this argument writes a copy of the
template code into a new R script.

```r
if (format == "plot") eval(code_output)
else if (format == "edit") {
  output <- glue(
    "# Explanation of how to edit the code here \n\n",
    "code_output$code <- ", "expr({{",
    system.file("extdata", "raincloud.R", package = "package name") |>
      readLines() |>
      paste(collapse = "\n"),
    "}})"
  )
  rstudioapi::documentNew(
    text = output,
    type = "r"
  )
  assign(
    "code_output",
    list(code = code_output, env = environment()),
    envir = .GlobalEnv
  )
}
```

This creates the editable script for the user and stores both the plot code and the wrapper's
environment in `code_output`.
The stored environment is important because it lets the edited code use objects created inside the
wrapper function, such as processed versions of the input data.

The original method is intended for a one-off change.
The user first runs the plot in its default form, but finds that it's hard to see the points with the
boxplot on top of them.
To make an adjustment they can rerun the function with format set to `"edit"`. 

\begin{fadecode}
\color[gray]raincloud_original(x = a)
\color[black]raincloud_original(x = a, format = "edit")
\end{fadecode}

Running `raincloud_original(x = a, format = "edit")` opens an unnamed R script containing the following code.
To make the points clearer, the user swaps the order of the jitter and boxplot layers. 

\begin{fadecode}
\color[black]code_output\$code <- expr({print(
\color[gray]  ggplot(x, aes(x = class, y = value, fill = class, color = class))+
\color[gray]    geom_violin(alpha = 0.25, trim = FALSE, width = 0.85, color = NA)+
\color[black]    geom_boxplot(width = 0.16, outlier.shape = NA, fill = "white", color = "black")+
\color[black]    geom_jitter(width = 0.12, alpha = 0.35, size = 1.2)+
\color[gray]    stat_summary(fun = median, geom = "point", size = 2.2, color = "black")+
\color[gray]    labs(x = NULL, y = NULL)+
\color[gray]    theme_minimal()+
\color[gray]    theme(legend.position = "none")
\color[gray]})
\end{fadecode}

After making the change, the user runs the edited block once to save the revised code object.
Then they call `raincloud_adjust()` to evaluate the modified version in the original function
environment.

\begin{fadecode}
\color[black]raincloud_adjust()
\end{fadecode}

```{r}
#| label: raincloud-original-adjusted
#| fig-cap: "Original-method edit: the boxplot is drawn behind the jittered points."
#| echo: false
ggplot(a, aes(x = class, y = value, fill = class, color = class))+
  geom_violin(alpha = 0.25, trim = FALSE, width = 0.85, color = NA)+
  geom_boxplot(width = 0.16, outlier.shape = NA, fill = "white", color = "black")+
  geom_jitter(width = 0.12, alpha = 0.35, size = 1.2)+
  stat_summary(fun = median, geom = "point", size = 2.2, color = "black")+
  labs(x = NULL, y = NULL)+
  theme_minimal()+
  theme(legend.position = "none")
```

This method is simple, but the modified code lives in the default `code_output` object.
If the user wants several different edited versions, or wants to name a particular version, the next
method is more convenient.

## Binding method

The binding method keeps the same basic edit branch, but returns an object to be
assigned as opposed to just saving to the global environment.
As a slight improvement a bnd class was created with a print method
that takes the code of the object and evaluates it in the
stored environment.

```r
if (format == "plot") eval(code_output)
else if (format == "edit") {
  output <- glue(
    "# Fill in the object name to the left of $code \n\n",
    "$code <- ", "expr({{",
    system.file("extdata", "raincloud.R", package = "package name") |>
      readLines() |>
      paste(collapse = "\n"),
    "}})"
  )
  rstudioapi::documentNew(
    text = output,
    type = "r"
  )
  bnd(list(code = code_output, env = environment()))
}
```

This means the user can save the returned object under a meaningful name, edit its `$code` slot, and
then print the object to evaluate the modified code in the original function environment.

The user finds that points are a little clearer now that they are on top, but that the color blends
into the violin plot background.
They decide they would like to change the shape of the points and make them black so they stand out
better.
In case they want to make further adjustments to the code, they use the binding method and assign it
to `black_points` in the global environment. 

\begin{fadecode}
\color[gray]raincloud_binding(x = a)
\color[black]black_points <- raincloud_binding(x = a, format = "edit")
\end{fadecode}

The editable script again contains the plot code, but the wrapper returns an object with class `bnd`.
The user fills in the object name on the left side of `$code`, edits the plot code, and runs the
block.

\begin{fadecode}
\color[black]black_points\$code <- expr({print(
\color[gray]  ggplot(x, aes(x = class, y = value, fill = class, color = class))+
\color[gray]    geom_violin(alpha = 0.25, trim = FALSE, width = 0.85, color = NA)+
\color[black]    geom_boxplot(width = 0.16, outlier.shape = NA, fill = "white", color = "black")+
\color[black]    geom_jitter(width = 0.12, alpha = 0.35, shape = 5, color = "black")+
\color[gray]    stat_summary(fun = median, geom = "point", size = 2.2, color = "black")+
\color[gray]    labs(x = NULL, y = NULL)+
\color[gray]    theme_minimal()+
\color[gray]    theme(legend.position = "none")
\color[gray]})
\end{fadecode}

Because `black_points` has a print method, the user can display the modified plot by evaluating the object itself.

\begin{fadecode}
\color[black]black_points
\end{fadecode}

```{r}
#| label: raincloud-binding-adjusted
#| fig-cap: "Binding-method edit: jittered points are drawn on top as black diamonds."
#| echo: false
ggplot(a, aes(x = class, y = value, fill = class, color = class))+
  geom_violin(alpha = 0.25, trim = FALSE, width = 0.85, color = NA)+
  geom_boxplot(width = 0.16, outlier.shape = NA, fill = "white", color = "black")+
  geom_jitter(width = 0.12, alpha = 0.35, shape = 5, color = "black")+
  stat_summary(fun = median, geom = "point", size = 2.2, color = "black")+
  labs(x = NULL, y = NULL)+
  theme_minimal()+
  theme(legend.position = "none")
```

The adjusted version now has a name in the user's workspace, so they won't have to worry about
overwriting it if they make other variants of this wrapper function.
The limitation is that it still lives inside the current R session unless the user saves it separately.

## Template method

For the template method, the creator uses an additional function that can create new templates for every wrapper function in the package. 

```r
make_template <- function(fun = NULL, template_name = "basic") {
  if (fun %in% c("raincloud")) {
    src <- system.file("extdata", paste0(fun, ".R"), package = "package name")
    dest <- paste0(template_name, ".R")
    file.copy(src, dest, overwrite = FALSE)
    invisible(rstudioapi::navigateToFile(dest))
  }
}
```

The wrapper itself then needs only a small branch that runs the normal plot when no template is
provided, or sources the user's edited template when one is provided:

```r
raincloud_template <- function(x, template = NULL) {
  if (is.null(template)) {
    # default plotting code
  } else {
    source(template, local = TRUE)
  }
}
```

The template file contains the plot code that the user is meant to edit.
Because the template is sourced locally, it can still use objects from inside the wrapper function.

The template method makes the edited version more durable.
Instead of storing the changed code only in the current environment, the user creates an editable `.R` template file.
The user decides that they would like to make one more adjustment, but decide to store it as a template for sharing.

\begin{fadecode}
\color[black]make_template("raincloud", "clearer_points")
\end{fadecode}

This copies the package's raincloud plotting template into `clearer_points.R` and opens it for
editing.
The user can then apply the previous two changes and make one additional adjustment: increasing the
jitter width from `0.12` to `0.35`, giving the diamond points more horizontal room to spread out.

\begin{fadecode}
\color[gray]print(
\color[gray]  ggplot(x, aes(x = class, y = value, fill = class, color = class))+
\color[gray]    geom_violin(alpha = 0.25, trim = FALSE, width = 0.85, color = NA)+
\color[black]    geom_boxplot(width = 0.16, outlier.shape = NA, fill = "white", color = "black")+
\color[black]    geom_jitter(width = 0.35, alpha = 0.35, shape = 5, color = "black")+
\color[gray]    stat_summary(fun = median, geom = "point", size = 2.2, color = "black")+
\color[gray]    labs(x = NULL, y = NULL)+
\color[gray]    theme_minimal()+
\color[gray]    theme(legend.position = "none")
\color[gray])
\end{fadecode}

Once the file is saved, the wrapper can use the modified template directly:

\begin{fadecode}
\color[black]raincloud_template(a, "clearer_points.R")
\end{fadecode}

```{r}
#| label: raincloud-template-adjusted
#| fig-cap: "Template-method edit: previous edits plus wider jitter spacing."
#| echo: false
ggplot(a, aes(x = class, y = value, fill = class, color = class))+
  geom_violin(alpha = 0.25, trim = FALSE, width = 0.85, color = NA)+
  geom_boxplot(width = 0.16, outlier.shape = NA, fill = "white", color = "black")+
  geom_jitter(width = 0.35, alpha = 0.35, shape = 5, color = "black")+
  stat_summary(fun = median, geom = "point", size = 2.2, color = "black")+
  labs(x = NULL, y = NULL)+
  theme_minimal()+
  theme(legend.position = "none")
```

This final version is the easiest to reuse across sessions or share with another user. 

# Conclusion

Considering the pros and cons of the methods described, we would have to recommend the template
method as being the most valuable for users.
The ability to not just adjust a function, but save those adjustments in an easily reusable and
shareable way, is valuable in any collaborative setting where wrapper functions are being used.
One can easily imagine creating plot templates with colors adjusted to match a company or school
theme and sharing them with colleagues.
The difficulty of implementation is also not particularly higher than other simpler methods
described.
The largest barrier to entry for all methods described remains how self-contained a wrapper function
is able to be, since code can only be edited if it is accessible to users. 

# References

::: {#refs}
:::
