---
format: pdf
editor_options: 
  chunk_output_type: console
---

# 25-12-05_update

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
theme_set(theme_minimal())
theme_update(panel.grid.minor = element_blank())
knitr::opts_chunk$set(warning = FALSE)
library(ggclassify)
library(simplex)
library(tidyclust)
library(tidymodels)
library(patchwork)

set.seed(17)
```

# Default Clustering behavior

The available k means engines available with tidyclust require picking
some number of clusters so the previous default was a rather arbitrary k = 3
k means model if the user didn't supply a method. 

To get a stronger default the new behavior if no method is supplied is 5 fold
cross validation across k = 2 to 10, where the k with the highest average silhouette
value is selected. The user is warned this was selected and that value is used
to plot. 

This does require tidymodels to use. 

```{r, fig.width=8}
samp <- rnorm_mixture(100, mean = roots_of_unity(5), var = .1)
df <- data.frame(x = samp[,1], y = samp[,2], c = attr(samp, "component"))
ggplot(df, aes(x, y))+
  geom_cluster()+
  geom_point(aes(col = c))
```

The model is pretty effective with Roots of Unity

```{r, fig.width=8}
set.seed(123)
means <- cbind(x = runif(6,-1), y = runif(6,-1))
samp <- rnorm_mixture(149, mean = means, var = .025)
bidata <- tibble(x = samp[,1], y = samp[,2], c = attr(samp, "component"))
ggplot(bidata, aes(x, y))+
  geom_cluster()+
  geom_point(aes(col = c))
```

More complex data can be difficult though.


# Documentation work

## Man pages

I have condensed the documentation from `geom_classify`, `geom_classify_boundary`, and `geom_point_classify` into one page. as well as `geom_cluster`, `geom_cluster_boundary`, and `geom_point_cluster` into one page. This should help reduce the amount of copy paste documentation. I've further outlined the computed stats available to each geom. 
