---
format: pdf
editor_options: 
  chunk_output_type: console
---

# 25-11-12_update

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
theme_set(theme_minimal())
theme_update(panel.grid.minor = element_blank())
knitr::opts_chunk$set(warning = FALSE)
library(ggclassify)
library(simplex)
library(tidyclust)
library(patchwork)

set.seed(17)

means <- cbind(x = runif(5,-1), y = runif(5,-1))
samp <- rnorm_mixture(100, mean = means, var = .1)
bidata <- tibble(x = samp[,1], y = samp[,2], class = attr(samp, "component"))
simpleplot <- ggplot(bidata, aes(x, y, fill = class))+
  geom_point_fill()
```

## Placing classes with non continuous color pallets next to each other

```{r, fig.width=8}
classify_boot_blend(bidata, simpleplot, n = 100, res = 101, method = "qda")+
  theme(legend.position = "none")+
ggplot(bidata, aes(x = x, y = y, fill = class))+
  geom_classify(method = "qda")+
  geom_point_fill() + plot_layout(guides = "collect", axes = "collect")
```

Same plot with KNN

```{r, fig.width=8, echo=FALSE}
classify_boot_blend(bidata, simpleplot, n = 100, res = 101, method = "knn")+
  theme(legend.position = "none")+

ggplot(bidata, aes(x = x, y = y, fill = class))+
  geom_classify(method = "knn")+
  geom_point_fill()+
  plot_layout(guides = "collect", axes = "collect")
```
\newpage
## Repeating those plots using the binned proportion of the plurality class.

```{r, fig.width=8}
classify_boot_blend(bidata, simpleplot, n = 100, res = 101, method = "qda", style = "fade")+
  
ggplot(bidata, aes(x = x, y = y, fill = class))+
  geom_classify(method = "qda")+
  theme(legend.position = "none")+
  geom_point_fill() + plot_layout(guides = "collect", axes = "collect")
```

```{r, fig.width=8, echo=FALSE}
classify_boot_blend(bidata, simpleplot, n = 100, res = 101, method = "knn", style = "fade")+

ggplot(bidata, aes(x = x, y = y, fill = class))+
  geom_classify(method = "knn")+
  geom_point_fill()+
  theme(legend.position = "none")+
  plot_layout(guides = "collect", axes = "collect")
```
\newpage

## geom_point_cluster

```{r, fig.width=8, fig.height=6}
kmeans_spec_5 <- k_means(num_clusters = 5) |> set_engine("stats")

(ggplot(bidata, aes(x, y))+
  geom_point_cluster(method = kmeans_spec_5))/
(ggplot(bidata, aes(x, y))+
  geom_point_fill(aes(fill = class)))
```

\newpage

## geom_point_classify

```{r, fig.width=8}
ggplot(bidata, aes(x, y, fill = class))+
  geom_point_classify()
```

color can be used instead of fill via the class afterstat

```{r, fig.width=8}
ggplot(bidata, aes(x, y, fill = class)) + 
  stat_point_classify(geom = GeomPoint, aes(col = after_stat(class)), shape = 19)
```

\newpage

An interesting way to see where the classifier fails is combining color and fill, or color and shape

```{r, fig.width=8}
ggplot(penguins, aes(bill_len, bill_dep, fill = species, col = species)) + 
  geom_point_classify()
```

```{r, fig.width=8}
ggplot(penguins, aes(bill_len, bill_dep, fill = species)) + 
  stat_point_classify(geom = GeomPoint,
                      aes(col = after_stat(class), shape = species))
```

\newpage

## Returning to a class aesthetic?

I went back through my previous implementation and was reminded of part of the reason
we moved on from class as an aesthetic. It didn't properly produce a legend. The implementation
looked like this.

```{r, eval=FALSE}
Statclassify <- ggproto("Statclassify", Stat,
                            setup_params = cpt_par,
                            setup_data = cpt_data,
                            compute_panel = cpt_class,
                            required_aes = c("x","y","class"),
                            default_aes = aes(fill = ..fill..)
)
```

I took in a class, but then internally computed a fill after stat and returned it. This plotted
fine, but the legend didn't know to link class to fill. I tried to experiment with allowing for class and fill to both be supplied for `geom_point_classify`, but while i got it to plot fine, the legend wouldn't appear for class aesthetic. 

```{r, include=FALSE}

cpt_class <- function(data, scales, method, args=NULL, 
                      mod = NULL, param = NULL) {
  if (!"fill" %in% names(data) && "class" %in% names(data)) {
    data$fill <- ggdistcol(length(levels(data$class)))[as.numeric(data$class)]
  }
  data <- do.call(get_classify,
                  args = c(list(data = data, mesh = data, method = method,
                                args = args), param))
  if (!"class" %in% names(data)) cbind(data, class = data$fill) else data
}




cpt_par = function(data, params) {
  #Default to LDA if nothing is given
  if( is.null(params$method) ){
    params$method <- "lda"
  }
  return(params)
}





cpt_data <- function(data, params) {
  return(data)
}

#' @rdname geom_point_classify
#' @export
StatPointClassify <- ggproto("StatPointClassify", Stat,
                        setup_params = cpt_par,
                        setup_data = cpt_data,
                        compute_panel = cpt_class,
                        required_aes = c("x", "y"),
                        computed_vars = c("class", "fill")
)





#' @rdname geom_point_classify
#' @export
stat_point_classify <- function(mapping = NULL, data = NULL, geom = GeomPointClassify,
                          position = "identity", na.rm = FALSE, show.legend = NA, 
                          inherit.aes = TRUE, method = NULL, param = NULL, ...) {
  layer(
    stat = StatPointClassify, data = data, mapping = mapping, geom = geom, 
    position = position, show.legend = show.legend, inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, method = method, param = param, ...)    
  )
}




#' @rdname geom_point_classify
#' @export
GeomPointClassify <- ggproto("GeomPointClassify", GeomPoint,
                        default_aes = aes(
                          shape = 21,
                          colour = "black",
                          size = 2,
                          alpha = NA,
                          stroke = .5,
                          fill = "transparent"
                        )
)





#' @rdname geom_point_classify
#' @export
geom_point_classify <- function(mapping = NULL, data = NULL, stat = StatPointClassify,
                          position = "identity", na.rm = FALSE, 
                          show.legend = NA, inherit.aes = TRUE, ...) {
  ggplot2::layer(
    geom = GeomPointClassify, mapping = mapping,  
    data = data, stat = stat, position = position, 
    show.legend = show.legend, inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```

## Is kmean working?

Having read further on kmeans one of the two engines used is from the stats
package of base R, so results should be the same. However their are a number of settings
available to change that could improve performance. 

### comparing base and tidyclust

```{r, fig.width=8}
kmeans_spec_5 <- k_means(num_clusters = 5) |> set_engine("stats")

x <- rbind(matrix(rnorm(100, sd = 0.3), ncol = 2),
           matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2))
colnames(x) <- c("x", "y")

# Everything below this I reran before each plot. 
cl <- kmeans(x, 5)
cl$centers
mod <- fit(kmeans_spec_5, ~x+y, data=data.frame(x))
df <- tibble(y = x[,2], x = x[,1], cluster = factor(extract_cluster_assignment(mod)$.cluster))

ggplot(df, aes(x,y))+
  geom_point(shape = 21, aes(col = cluster))+
  theme(legend.position = "none")+
~plot(x, col = cl$cluster)
dev.off()
```

### Further repetition

```{r, fig.width=8, fig.height=3, echo=FALSE}
cl <- kmeans(x, 5)
mod <- fit(kmeans_spec_5, ~x+y, data=data.frame(x))
df <- tibble(y = x[,2], x = x[,1], cluster = factor(extract_cluster_assignment(mod)$.cluster))

ggplot(df, aes(x,y))+
  geom_point(shape = 21, aes(col = cluster))+
  theme(legend.position = "none")+
~plot(x, col = cl$cluster)
```
```{r, fig.width=8, fig.height=3, echo=FALSE}
cl <- kmeans(x, 5)
mod <- fit(kmeans_spec_5, ~x+y, data=data.frame(x))
df <- tibble(y = x[,2], x = x[,1], cluster = factor(extract_cluster_assignment(mod)$.cluster))

ggplot(df, aes(x,y))+
  geom_point(shape = 21, aes(col = cluster))+
  theme(legend.position = "none")+
~plot(x, col = cl$cluster)
```
```{r, fig.width=8, fig.height=3, echo=FALSE}
cl <- kmeans(x, 5)
mod <- fit(kmeans_spec_5, ~x+y, data=data.frame(x))
df <- tibble(y = x[,2], x = x[,1], cluster = factor(extract_cluster_assignment(mod)$.cluster))

ggplot(df, aes(x,y))+
  geom_point(shape = 21, aes(col = cluster))+
  theme(legend.position = "none")+
~plot(x, col = cl$cluster)
```

Across all these plots there are a couple variations in results for both implementations for tidyclust. They do appear to work the same

### improving performance

It turns out the default parameters for kmeans are simply terrible, just by changing
the number of starts to 10 instead of the default 1 performance seems much better

```{r}
kmeans_spec_5_improved <- k_means(num_clusters = 5) |> set_engine("stats", nstart = 10)
```


```{r, fig.width=8, fig.height=3, echo=FALSE}
ggplot(df, aes(x,y)) + 
  geom_cluster(method = kmeans_spec_5) + 
  geom_point(shape = 21, aes(fill = cluster))+
  ggtitle("old")+
ggplot(df, aes(x,y)) + 
  geom_cluster(method = kmeans_spec_5_improved) + 
  geom_point(shape = 21, aes(fill = cluster))+
  ggtitle("improved")+
  plot_layout(guides = "collect", axes = "collect")

ggplot(df, aes(x,y)) + 
  geom_cluster(method = kmeans_spec_5) + 
  geom_point(shape = 21, aes(fill = cluster))+
  ggtitle("old")+
ggplot(df, aes(x,y)) + 
  geom_cluster(method = kmeans_spec_5_improved) + 
  geom_point(shape = 21, aes(fill = cluster))+
  ggtitle("improved")+
  plot_layout(guides = "collect", axes = "collect")

ggplot(df, aes(x,y)) + 
  geom_cluster(method = kmeans_spec_5) + 
  geom_point(shape = 21, aes(fill = cluster))+
  ggtitle("old")+
ggplot(df, aes(x,y)) + 
  geom_cluster(method = kmeans_spec_5_improved) + 
  geom_point(shape = 21, aes(fill = cluster))+
  ggtitle("improved")+
  plot_layout(guides = "collect", axes = "collect")

ggplot(df, aes(x,y)) + 
  geom_cluster(method = kmeans_spec_5) + 
  geom_point(shape = 21, aes(fill = cluster))+
  ggtitle("old")+
ggplot(df, aes(x,y)) + 
  geom_cluster(method = kmeans_spec_5_improved) + 
  geom_point(shape = 21, aes(fill = cluster))+
  ggtitle("improved")+
  plot_layout(guides = "collect", axes = "collect")

ggplot(df, aes(x,y)) + 
  geom_cluster(method = kmeans_spec_5) + 
  geom_point(shape = 21, aes(fill = cluster))+
  ggtitle("old")+
ggplot(df, aes(x,y)) + 
  geom_cluster(method = kmeans_spec_5_improved) + 
  geom_point(shape = 21, aes(fill = cluster))+
  ggtitle("improved")+
  plot_layout(guides = "collect", axes = "collect")
```



