---
title: "25-10-22_update"
format: pdf
editor_options: 
  chunk_output_type: console
---

# Update to make_template creating a file and opening into new window

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
theme_set(theme_minimal())
theme_update(panel.grid.minor = element_blank())
knitr::opts_chunk$set(warning = FALSE)
library(ggclassify)
library(simplex)

set.seed(17)

df <- data.frame(x = rnorm(40), y=rnorm(40) + 2, class = factor("A")) |> 
  rbind(data.frame(x = rnorm(40, sd=2) + 2, y=rnorm(40, sd = 2) - 1, class = factor("B"))) |> 
  rbind(data.frame(x = rnorm(40, sd=1/2) + 2, y=rnorm(40, sd = 1/2) + 1,class = factor("C")))
```

```{r}
#Make Data
set.seed(17)
a <- tibble(value = c(rnorm(2e2, 1), 4*rbeta(2e2, 1, 4)),
            class = c(rep("a", 2e2),rep("b", 2e2)))
boxplot_template(a)


#Create Blue template that has a blue fill
#make_template("boxplot", "blue")
read_lines(file = "blue.R")


#make the boxplot with the new template
boxplot_template(a, "blue.R")
```

# Create wrapper functions in ggclassify

## Boundary bootwrapping function

```{r}
simpleplot <- ggplot(df, aes(x, y, fill = class))+
   geom_point_fill()

classify_boundary_boot(df, simpleplot, n = 100, method = "lda")
classify_boundary_boot(df, simpleplot, n = 100, method = "qda")
classify_boundary_boot(df, simpleplot, n = 100, method = "knn")
classify_boundary_boot(df, simpleplot, n = 100, method = "randomForest")
```

## Piechart style bootwrapping function

The current issue is that this wrapper relies upon a geom from another package with multiple dependencies.

```{r}
classify_boot(df, simpleplot, n = 100, res = 21, method = "lda")
classify_boot(df, simpleplot, n = 100, res = 21, method = "qda")
classify_boot(df, simpleplot, n = 100, res = 21, method = "knn")
classify_boot(df, simpleplot, n = 100, res = 21, method = "randomForest")
```

# Non-progressed

## Geom_Cluster
Have not yet completed rewriting tidyclust implementation into separate geom_cluster

## Potrace


\newpage
# Continuing work on the outline for wrapper function paper, looks something like this

```{r}
# Wrapper Functions Background

## What are wrapper functions
### Functions that try to make a complex process easy to run 
### ggplot wrapper functions (the focus of this paper?)

## Advantages to wrappers
### Complex analysis and plots can be created in one line by new user
### examples of popular wrappers

## How they can obscure user choice
### plot and function parameters that arn't adjustable
### example of wrapper plots that can be difficult to edit (time series comes to mind)
### Complex crash messages that arn't clear about their source

# Issues with developing wrapper functions

## The balance between simplicity for user, and ability to customize.
### More options for input can lead to increasing code complexity when trying to account for every combination

## Self contained versus non self contained wrapper functions?
### Functions that call other functions
#### Functions that call non-exported functions
#### Functions that call C code. 
### Functions that utilize S3 or other methods. 

# Potential solutions to simplicity & Customization tradeoff

## "original" method (Perhaps so simple that I just combine it and the name method which have very little difference)
### The simplest default solution I originally developed
### Runs into issues if making more than 1 or 2 custom plots

## name method
### A slight improvement on that in the form of the "name" method
### Alows for storing multiple different customized plots since each has seperate binding name

## binding method
### Can utlize S3 print methods and is more explicit
### Doesn't save lists into user global without their knowlege

## template method
### Uses .R files to store custom made templates for further use

## Why i think the template method is the best
### Templates can easily be shared between various users or stored between sessions
### Easier to further edit an R file after saving as opposed to a list stored in your global

# Utilizing these solutionson pre - existing wrappers

## Simple method for editing other functions
### basic function that can pull code of a function into a new window for editing and saving a new version of that wrapper function
### Will run into issues on any non-self contained wrapper function (many of them)
### The best implementations will involve the author including these methods of editing themselves

## Not sure but If i try to make a roxygen tag it goes here

## how wrapper writers could utlize these potential solutions (This might should be done in the previous section as I explain the methods)
```

