---
format: pdf
editor_options: 
  chunk_output_type: console
---

# 25-11-12_update

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
theme_set(theme_minimal())
theme_update(panel.grid.minor = element_blank())
knitr::opts_chunk$set(warning = FALSE)
library(ggclassify)
library(simplex)
library(tidyclust)
library(patchwork)

set.seed(17)

means <- cbind(x = runif(5,-1), y = runif(5,-1))
samp <- rnorm_mixture(100, mean = means, var = .1)
bidata <- tibble(x = samp[,1], y = samp[,2], class = attr(samp, "component"))
simpleplot <- ggplot(bidata, aes(x, y, fill = class))+
  geom_point_fill()
```

## Placing classes with non continuous color pallets next to each other

```{r, fig.width=8}
classify_boot_blend(bidata, simpleplot, n = 100, res = 101, method = "qda")+
  theme(legend.position = "none")+
ggplot(bidata, aes(x = x, y = y, fill = class))+
  geom_classify(method = "qda")+
  geom_point_fill() + plot_layout(guides = "collect", axes = "collect")
```

Same plot with KNN

```{r, fig.width=8, echo=FALSE}
classify_boot_blend(bidata, simpleplot, n = 100, res = 101, method = "knn")+
  theme(legend.position = "none")+

ggplot(bidata, aes(x = x, y = y, fill = class))+
  geom_classify(method = "knn")+
  geom_point_fill()+
  plot_layout(guides = "collect", axes = "collect")
```
\newpage
## Repeating those plots using the binned proportion of the plurality class.

```{r, fig.width=8}
classify_boot_blend(bidata, simpleplot, n = 100, res = 101, method = "qda", style = "fade")+
  
ggplot(bidata, aes(x = x, y = y, fill = class))+
  geom_classify(method = "qda")+
  theme(legend.position = "none")+
  geom_point_fill() + plot_layout(guides = "collect", axes = "collect")
```

```{r, fig.width=8, echo=FALSE}
classify_boot_blend(bidata, simpleplot, n = 100, res = 101, method = "knn", style = "fade")+

ggplot(bidata, aes(x = x, y = y, fill = class))+
  geom_classify(method = "knn")+
  geom_point_fill()+
  theme(legend.position = "none")+
  plot_layout(guides = "collect", axes = "collect")
```
\newpage

## geom_point_cluster

```{r, fig.width=8, fig.height=6}
kmeans_spec_5 <- k_means(num_clusters = 5) |> set_engine("stats")

(ggplot(bidata, aes(x, y))+
  geom_point_cluster(method = kmeans_spec_5))/
(ggplot(bidata, aes(x, y))+
  geom_point_fill(aes(fill = class)))
```

\newpage

## geom_point_classify

```{r, fig.width=8}
ggplot(bidata, aes(x, y, fill = class))+
  geom_point_classify()
```

color can be used instead of fill via the class afterstat

```{r, fig.width=8}
ggplot(bidata, aes(x, y, fill = class)) + 
  stat_point_classify(geom = GeomPoint, aes(col = after_stat(class)), shape = 19)
```

\newpage

An interesting way to see where the classifier fails is combining color and fill, or color and shape

```{r, fig.width=8}
ggplot(penguins, aes(bill_len, bill_dep, fill = species, col = species)) + 
  geom_point_classify()
```

```{r, fig.width=8}
ggplot(penguins, aes(bill_len, bill_dep, fill = species)) + 
  stat_point_classify(geom = GeomPoint,
                      aes(col = after_stat(class), shape = species))
```
