---
format: pdf
editor_options: 
  chunk_output_type: console
---

# 25-12-05_update

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
theme_set(theme_minimal())
theme_update(panel.grid.minor = element_blank())
knitr::opts_chunk$set(warning = FALSE)
library(ggclassify)
library(simplex)
library(tidyclust)
library(tidymodels)
library(patchwork)

set.seed(17)
```

# Default Clustering behavior

The available k means engines available with tidyclust require picking
some number of clusters so the previous default was a rather arbitrary k = 3
k means model if the user didn't supply a method. 

To get a stronger default the new behavior if no method is supplied is 5 fold
cross validation across k = 2 to 10, where the k with the highest average silhouette
value is selected. The user is warned this was selected and that value is used
to plot. 

This does require tidymodels to use. 

```{r, fig.width=8}
samp <- rnorm_mixture(100, mean = roots_of_unity(5), var = .1)
df <- data.frame(x = samp[,1], y = samp[,2], c = attr(samp, "component"))
ggplot(df, aes(x, y))+
  geom_cluster()+
  geom_point(aes(col = c))
```

The model is pretty effective with Roots of Unity

```{r, fig.width=8}
set.seed(123)
means <- cbind(x = runif(6,-1), y = runif(6,-1))
samp <- rnorm_mixture(149, mean = means, var = .025)
bidata <- tibble(x = samp[,1], y = samp[,2], c = attr(samp, "component"))
ggplot(bidata, aes(x, y))+
  geom_cluster()+
  geom_point(aes(col = c))
```

More complex data can be difficult though.

I have my method for selecting k below. Not the most experience with
building tidymodels so their could be inefficiencies. 

```{r, eval=FALSE}
get_best_kmeans <- function(data){
  
  data_cv <- vfold_cv(data, v = 5)
  
  data_rec <- step_normalize(recipe(~ x + y, data = data), all_predictors())
  
  kmeans_spec_tune <- set_engine(k_means(num_clusters = tune()), "stats", iter.max = 25, nstart = 3)
  
  kmeans_wflow <- add_model(add_recipe(workflow(), data_rec), kmeans_spec_tune)
  
  clust_metrics <- cluster_metric_set(silhouette_avg)
  
  data_tune <- tune_cluster(
    kmeans_wflow,
    resamples = data_cv,
    grid = tibble(num_clusters = 2:10),
    metrics = clust_metrics
  )
  
  metrics_tbl <- collect_metrics(data_tune)

  best_k <- metrics_tbl$num_clusters[which.max(metrics_tbl$mean)]
  cli_alert_warning(paste0("a k means model with ", best_k ," clusters was used"))
  set_engine(k_means(num_clusters = best_k), "stats", iter.max = 40, nstart = 10) 
}
```


```{r}
datasets::penguins |> remove_missing() |> mutate(x = bill_len, y = bill_dep) |> 
  ggclassify:::get_best_kmeans()
```

# Documentation work

## Man pages

I have condensed the documentation from `geom_classify`, `geom_classify_boundary`, and `geom_point_classify` into one page. as well as `geom_cluster`, `geom_cluster_boundary`, and `geom_point_cluster` into one page. This should help reduce the amount of copy paste documentation. I've further outlined the computed stats available to each geom. 
